Appendix S2: Data Exploration

### 2021-09-06 ms/data_exploration.Rmd
Make data exploration a pdf (SI2)

---
title: "Data exploration"
output: html_document
editor_options: 
  chunk_output_type: console
params:
  knit_type: manual
---

@@

```{r load-packages, include = params$knit_type == "manual"}
# Load packages
library(tidyverse)
library(targets)
library(outliers)
library(ggforce)
library(assertr)

# Load custom functions
source(here::here("R/functions.R"))
```

The dataset includes biodiversity metrics and environmental data.

All biodiversity metrics were calculated during the `targets` workflow (`_targets.R`).

The `cent` part of the name indicates that they include geographic locations as centroids of each grid cell.

```{r load-data}
# Load the data from the targets workflow
tar_load(biodiv_ferns_spatial)
tar_load(resp_vars_env)
tar_load(indep_vars)

# Make biodiversity metrics dataframe with centroid of each site for models
  biodiv_ferns_cent <- 
    biodiv_ferns_spatial %>%
    # keep only variables needed for model and only rows with zero missing data
    filter_data_for_model(c("grids", "lat", "long", resp_vars_env, indep_vars)) %>%
    sf::st_set_geometry(NULL) %>%
    as_tibble()
```

@@

```{r struc-full-data}
biodiv_ferns_cent
```

@@

* `grids`: ID code for each grid cell
* `lat`: latitude of grid cell centroid
* `long`: longitude of grid cell centroid
* `richness`: number of taxa
* `pd_obs_z`: standard effect size (SES) of phylogenetic diversity (pd)
* `fd_obs_z`: SES of functional diversity (fd)
* `rpd_obs_z`: SES of relative pd
* `rfd_obs_z`: SES of relative fd
* `percent_apo`: percent of apogamous taxa in that grid cell
* `temp`: mean annual temperature (10 x °C)
* `precip`: total annual precipitation (mm)
* `precip_season`: seasonality in precipitation
* `area`: rolling mean of area in 0.2 degree latitudinal bands

@@

```{r check-missing}
# These will return TRUE if all data are non-missing, 
# or an error if any data are missing (NA)
biodiv_ferns_cent %>%
  assert(not_na, everything(), success_fun = success_logical)
```

No missing data.

@@

```{r box-plots-full}
biodiv_ferns_cent %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot() +
  facet_wrap(vars(variable), scales = "free") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank()
  )
```

There are a few points beyond the whiskers that may be outliers.

Next, check for outliers using Grubb's test. Significant p-values indicate that the most extreme value is an outlier.

```{r grubbs-env}
biodiv_ferns_cent %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  group_by(variable) %>%
  nest() %>%
  mutate(map_df(data, ~grubbs_tidy(.$value))) %>%
  select(-data) %>%
  ungroup() %>%
  select(variable:p_value) %>%
  mutate(signif = p_value < 0.05)
```

So the one percent apomictic > 60% appears to be an outlier. Perhaps this has a very small number of species, most of which happen to be apogamous?

```{r check-apo-outlier}
biodiv_ferns_cent %>%
  slice_max(1, order_by = percent_apo) %>%
  left_join(select(biodiv_ferns_cent, grids, richness)) %>%
  select(grids:long, percent_apo, richness)
```

Looks like that is the case---just three species in this grid cell! This data point should be excluded. Let's do that and check for outliers again.

```{r remove-apo-outlier-check-grubbs}
biodiv_ferns_cent %>% 
  filter(percent_apo != max(percent_apo)) %>%
  pull(percent_apo) %>%
  grubbs_tidy()
```

@@

Check for normality with histograms and QQ-plots.

Inspect histograms.

```{r hist-env}
biodiv_ferns_cent %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30) +
  facet_wrap(vars(variable), scales = "free")
```

Inspect qq-plots.

```{r qqplot-env}
biodiv_ferns_cent %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(sample = value)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~variable, scales = "free") +
  labs(title = "Full dataset")
```

@@

```{r check-for-zeros}
biodiv_ferns_cent %>%
  mutate(rich_zero = richness == 0) %>%
  arrange(richness) %>%
  summarize(
    total_zero = sum(rich_zero),
    perc_zero = sum(rich_zero) / nrow(.))
```

@@

Here we plot response variables (row labels) vs. indep variables (column labels) for each dataset, and add a trend line.

```{r plot-raw-data, fig.height = 8}
ggplot(biodiv_ferns_cent, aes(x = .panel_x, y = .panel_y)) + 
  geom_point(alpha = 0.5, shape = 16, size = 0.5) + 
  geom_smooth() +
  facet_matrix(
    cols = vars(temp, precip, precip_season, percent_apo),
    rows = vars(richness, pd_obs_z, fd_obs_z, rpd_obs_z, rfd_obs_z))
```

`richness` vs. `temp` definitely shows a hump-shaped trend line, and the other response variables vs. `temp` also show one if not quite as strongly. The other relationships appear to be linear or uncorrelated.

```{r plot-raw-data-repro}
# Remove the outlier first
biodiv_ferns_cent %>% 
  filter(percent_apo != max(percent_apo)) %>%
ggplot(aes(x = .panel_x, y = .panel_y)) + 
  geom_point(alpha = 0.5, shape = 16, size = 0.5) + 
  geom_smooth() +
  facet_matrix(
    cols = vars(percent_apo, temp, precip, precip_season),
    rows = vars(pd_obs_z, rpd_obs_z))
```

As one would expect from the last plot, `temp` shows a (negative) hump shape. However, there aren't many points on the right.. `percent_apo` seems more linear, especially for `rpd_obs_z`, but it's subtle. Anyways, we will probably keep the quadratic term for temperature but use a linear model for `percent_apo`.

@@

```{r check-overdispersion}
# Create general linear model with poisson distribution
richness_poisson_mod <- glm(
  richness ~ temp + I(temp^2) + precip + precip_season, 
  data = biodiv_ferns_cent, family = poisson)

# Check ratio of residual variance to df (should be near 1 to meet poisson assumptions)
richness_poisson_mod$deviance / richness_poisson_mod$df.residual
```

That's very overdispersed. So we won't use poisson, but negative binomial instead.

>>>>

---
title: "Appendix S2: Data Exploration"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: yes
    keep_tex: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
editor_options: 
  chunk_output_type: console
header-includes: 
  - \usepackage{float}
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
---

@@

All code is in R, and workflow managed with the `targets` package (https://github.com/ropensci/targets).

The code here is included in the repo at https://github.com/joelnitta/japan_ferns_biogeo.

Let's get started by loading packages and tidying our namespace.

```{r load-packages, message = FALSE}
# Load packages
library(tidyverse)
library(targets)
library(outliers)
library(ggforce)
library(assertr)
library(conflicted)

# Resolve possible namespace conflicts
conflict_prefer("map", "purrr")
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("gather", "tidyr")
conflict_prefer("extract", "magrittr")
```

The dataset includes biodiversity metrics and environmental data measured on native, non-hybrid ferns of Japan.

All biodiversity metrics were calculated during the `targets` workflow (`_targets.R`), and are contained in the `biodiv_ferns_cent_dryad` object in the `targets` cache.

The next step uses `tar_load()` to load the dataset from the targets cache. Alternatively, this can be loaded from the dryad data file as shown in the commented-out code below.

```{r load-data}
# Load the data from the targets workflow
tar_load(biodiv_ferns_cent_dryad)

# Alternatively, unzip `results.zip` in the Dryad data
# (FIXME: add DOI when available)
# and read in the data from there
# biodiv_ferns_cent_dryad <- read_csv("japan_ferns_biodiv.csv")
```

This dataframe actually has more variables than are needed for modeling, so we will subset it to only the relevant variables.

```{r subset-vars}
biodiv_ferns_cent_raw <- 
  biodiv_ferns_cent_dryad %>%
  select(
    grids, lat, long, # grid-cell ID, latitude, longitude
    richness, pd_obs_z, fd_obs_z, rpd_obs_z, rfd_obs_z, # response vars
    percent_apo, temp, precip, precip_season, lat_area  # independent vars
  )
```

@@

```{r struc-full-data}
biodiv_ferns_cent_raw
```

@@

* `grids`: ID code for each grid cell
* `lat`: latitude of grid cell centroid
* `long`: longitude of grid cell centroid
* `richness`: number of taxa
* `pd_obs_z`: standard effect size (SES) of phylogenetic diversity (pd)
* `fd_obs_z`: SES of functional diversity (fd)
* `rpd_obs_z`: SES of relative pd
* `rfd_obs_z`: SES of relative fd
* `percent_apo`: percent of apogamous taxa in that grid cell
* `temp`: mean annual temperature (10 x °C)
* `precip`: total annual precipitation (mm)
* `precip_season`: seasonality in precipitation
* `lat_area`: rolling mean of area in 0.2 degree latitudinal bands

@@

Next, check for missing data. If any rows are missing data, exclude them.

```{r check-missing}
# Run a logical check to see if any rows have missing data 
# Returns TRUE if there is no missing data, FALSE if there is missing data
biodiv_ferns_cent_raw %>%
  assert(
    not_na, everything(), 
    error_fun = error_logical, 
    success_fun = success_logical)

# At least one row has missing data. 
# Check which.
error_rows <-
biodiv_ferns_cent_raw %>%
  assert(not_na, everything(), error_fun = error_df_return) %>%
  pull(index) %>%
  unique()

biodiv_ferns_cent_raw %>% slice(error_rows)

# The missing data was due to one grid-cell that lacks environmental data.
# Assume this must have been missing from WorldClim dataset.
# Remove the row with missing data, then continue
biodiv_ferns_cent_no_missing <-
  biodiv_ferns_cent_raw %>%
  slice(-error_rows)

# Make sure the resulting dataframe has no missing data
assert(biodiv_ferns_cent_no_missing, 
       not_na, everything(), 
       success_fun = success_logical)
```

@@

```{r box-plots-full}
biodiv_ferns_cent_no_missing %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(x = variable, y = value)) +
  geom_boxplot() +
  facet_wrap(vars(variable), scales = "free") +
  theme(
    axis.title.x = element_blank(),
    axis.text.x = element_blank()
  )
```

There are a few points beyond the whiskers that may be outliers. In particular, there is one value of `percent_apo` > 60% that seems quite outside the normal range.

Next, check for outliers using Grubb's test. Significant *p*-values indicate that the most extreme value is an outlier.

```{r grubbs-env}
biodiv_ferns_cent_no_missing %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  group_by(variable) %>%
  nest() %>%
  mutate(map_df(data, ~grubbs_tidy(.$value))) %>%
  select(-data) %>%
  ungroup() %>%
  select(variable:p_value) %>%
  mutate(signif = p_value < 0.05)
```

Grubb's test confirms that the percent apomictic value > 60% is an outlier. It also indicates that the highest values of the variables `fd_obs_z`, `pd_obs_z`, `rfd_obs_z`, and `rpd_obs_z` may be outliers.

One reason for artificially extreme values could be undersampling of species. Let's check richness of these possible outlier grid-cells.

```{r check-richness-apo}
biodiv_ferns_cent_no_missing %>%
  slice_max(1, order_by = percent_apo) %>%
  select(grids:long, percent_apo, richness)
```

There are just three species in the grid-cell with > 60% apomictic taxa. This data point should be excluded.

What about the other possible outliers?

```{r check-richness-other}
biodiv_ferns_cent_no_missing %>% 
  slice_max(n = 2, order_by = fd_obs_z) %>%
  select(grids:long, fd_obs_z, richness)

biodiv_ferns_cent_no_missing %>% 
  slice_max(n = 2, order_by = pd_obs_z) %>%
  select(grids:long, pd_obs_z, richness)

biodiv_ferns_cent_no_missing %>% 
  slice_max(n = 2, order_by = rfd_obs_z) %>%
  select(grids:long, rfd_obs_z, richness)

biodiv_ferns_cent_no_missing %>% 
  slice_max(n = 2, order_by = rpd_obs_z) %>%
  select(grids:long, rpd_obs_z, richness)
```

They all come from the same two grid-cells, but don't have an extremely low number of species.

So we will conservatively only remove the grid-cell with unusually high % apomictic taxa, then run Grubb's test again.

```{r remove-apo-outlier-check-grubbs}
biodiv_ferns_cent_no_outliers <-
  biodiv_ferns_cent_no_missing %>% 
  filter(percent_apo != max(percent_apo))

biodiv_ferns_cent_no_outliers %>%
  pull(percent_apo) %>%
  grubbs_tidy()
```

@@

Check for normality with histograms and QQ-plots after removing missing data and outliers.

First inspect histograms:

```{r hist-env}
biodiv_ferns_cent_no_outliers %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30) +
  facet_wrap(vars(variable), scales = "free")
```

Next inspect qq-plots:

```{r qqplot-env}
biodiv_ferns_cent_no_outliers %>%
  select(-lat, -long) %>%
  pivot_longer(-grids, names_to = "variable") %>%
  ggplot(aes(sample = value)) +
  stat_qq() +
  stat_qq_line() +
  facet_wrap(~variable, scales = "free")
```

@@

```{r check-for-zeros}
biodiv_ferns_cent_no_outliers %>%
  mutate(rich_zero = richness == 0) %>%
  arrange(richness) %>%
  summarize(
    total_zero = sum(rich_zero),
    perc_zero = sum(rich_zero) / nrow(.))
```

@@

Here we plot response variables (row labels) vs. indep variables (column labels) and add a trend line for each to get an idea of the possible model fits.

```{r plot-raw-data, fig.height = 8}
ggplot(biodiv_ferns_cent_no_outliers, aes(x = .panel_x, y = .panel_y)) + 
  geom_point(alpha = 0.5, shape = 16, size = 0.5) + 
  geom_smooth() +
  facet_matrix(
    cols = vars(temp, precip, precip_season, percent_apo),
    rows = vars(richness, pd_obs_z, fd_obs_z, rpd_obs_z, rfd_obs_z))
```

`richness` vs. `temp` definitely shows a hump-shaped trend line, and the other response variables vs. `temp` also show one if not quite as strongly. The other relationships appear to be linear or uncorrelated (the bend in the line for `percent_apo` vs. `richness` seems to be due to a relatively few number of points with > 30% apomictic taxa, but otherwise it looks linear).

@@

```{r check-overdispersion}
# Create general linear model with poisson distribution
richness_poisson_mod <- glm(
  richness ~ temp + I(temp^2) + precip + precip_season, 
  data = biodiv_ferns_cent_no_outliers, family = poisson)

# Check ratio of residual variance to df 
# (should be near 1 to meet poisson assumptions)
richness_poisson_mod$deviance / richness_poisson_mod$df.residual
```

That's very overdispersed. So we won't use poisson for `richness`, but negative binomial instead.

### Session information

```{r sesh}
sessionInfo()
```

### 2021-09-06 ms/data_readme.Rmd
Update SI
- Add plot for latitudinal patterns
- Fix lumping of bioregions in SI plot
Filter estimated deer range by movement cost < 0.10
Make data exploration a pdf (SI2)

# Load objects from targets workflow (all paths to data files)
tar_load(
  c(
    "ms_functions",
    "ja_climate_data_file",
    "japan_map_file",
    "deer_range_file",
    "japan_ferns_comm_full_file",
    "inext_res_file",
    "occ_point_data_summary_file",
    "plastome_calibration_dates",
    "redundancy_by_res_file",
    "japan_ferns_shape_full_file",
    "raw_trait_data_file",
    "japan_map_points_file",
    "protected_areas_file",
    "calibration_dates_file",
    "japan_ferns_biodiv_dryad_file",
    "comm_ferns_dryad_file",
    "shape_ferns_dryad_file",
    "fern_traits_dryad_file",
    "japan_fern_phylogram_dryad_file",
    "japan_fern_tree_dryad_file"
    )
  )
```

@@

- ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment and phylogenetic tree 
- japan_climate.gpkg: Climate data in Japan downloaded from WorldClim database
- japan_deer_range.gpkg: Distribution maps of Japanese deer (Cervus nippon) in Japan
- japan_ferns_comm_full.csv: Community matrix (species x sites matrix) of native, non-hybrid ferns in Japan, full (unfiltered) dataset
- japan_ferns_inext_results.csv: Results of collection curve analysis on fern specimens in Japan
- japan_ferns_occ_summary.csv: Summary of specimen data of native, non-hybrid ferns in Japan
- japan_ferns_redundancy_by_res: Taxonomic richness, abundance, and redunancy at different spatial resolutions for native, non-hybrid ferns of Japan
- japan_ferns_shape_full.gpkg: Location of grid-cells (sites) for native, non-hybrid ferns in Japan, full (unfiltered) dataset
- japan_ferns_traits_lucid.csv: Morphological trait data of native, non-hybrid ferns in Japan
- japan_map_points.csv: Latitude and longitude of points used to produce a labeled map of Japan
- japan_map.gpkg: Map of Japan
- japan_protected_areas.gpkg: Protected areas in Japan
- testo_sundue_2016_calibrations.csv: Fossil calibration points used for phylogenetic dating analysis
- results.zip: Selected analysis results. Files include:
  - japan_ferns_biodiv.csv: Biodiversity statistics of native, non-hybrid ferns in Japan
  - japan_ferns_comm.csv: Community matrix (species x sites matrix) of native, non-hybrid ferns in Japan used for biodiversity analysis
  - japan_ferns_shape.gpkg: Location of grid-cells (sites) for native, non-hybrid ferns in Japan used for biodiversity analysis
  - japan_ferns_traits.csv: Trait matrix of native, non-hybrid ferns in Japan used for functional biodiversity analysis
  - japan_ferns_tree_dated.tre: Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in Japan
  - japan_ferns_tree.tre: Maximim-likelihood phylogeny of native, non-hybrid ferns in Japan

@@

`r fs::path_file(deer_range_file)`: Distribution maps of Japanese deer (Cervus
nippon) in Japan. Raw data were downloaded from the Japan Ministry of the
Environment
(https://www.biodic.go.jp/biodiversity/activity/policy/map/map14/index.html).
The data include range maps based on three types of data: range of Japanese deer
surveyed in 1978, range surveyed in 2003, and estimated range inferred from a
model including snow cover and forest type based on the 2003 survey data. Only
estimated range with <0.05 movement cost was included.

@@

@@

```{r japan-ferns-biodiv-dryad}
japan_ferns_biodiv_dryad <- read_csv(here::here(japan_ferns_biodiv_dryad_file))
# Check that no data are missing
assert(japan_ferns_biodiv_dryad, not_na, everything(), success_fun = success_logical)
```

`r fs::path_file(japan_ferns_biodiv_dryad_file)` (contained in "results.zip"):
Biodiversity statistics of native, non-hybrid ferns and environmental variables
in Japan. Biodiversity metrics calculated as described in Nitta et al. 2021. A
single outlier grid-cell with extremely low taxon richness and extremely high %
apomictic taxa was excluded. Climatic (temperature and preciptation) variables
calculated as described for japan_climate.gpkg. Data used for spatial models
in Nitta et al. 2021.

@@

@@

`r fs::path_file(japan_fern_tree_dryad_file)` (contained in "results.zip"):
Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in
Japan in newick format. Tree dated using treePL v1.0 [@Smith2012] with `r
n_fossil_points` fossil calibration points after @Testo2016a. Values at nodes
are bootstrap support values calculated using 1000 replicates of ultrafast
bootstrap [@Nguyen2015].

@@

2021-09-03

>>>>

# Load objects from targets workflow (all paths to data files)
tar_load(
  c(
    "ms_functions",
    "ja_climate_data_file",
    "japan_map_file",
    "deer_range_file",
    "japan_ferns_comm_full_file",
    "inext_res_file",
    "occ_point_data_summary_file",
    "plastome_calibration_dates",
    "redundancy_by_res_file",
    "japan_ferns_shape_full_file",
    "raw_trait_data_file",
    "japan_map_points_file",
    "protected_areas_file",
    "calibration_dates_file",
    "japan_ferns_biodiv_dryad_file",
    "comm_ferns_dryad_file",
    "shape_ferns_dryad_file",
    "fern_traits_dryad_file",
    "japan_fern_phylogram_dryad_file",
    "japan_fern_tree_dryad_file",
    "japan_fern_tree_uncollapsed_dryad_file"
    )
  )
```

@@

- ftol_data_release_v0.0.1.zip: Fern Tree of Life project global DNA alignment and phylogenetic tree 
- japan_climate.gpkg: Climate data in Japan downloaded from WorldClim database
- japan_deer_range.gpkg: Distribution maps of Japanese deer (Cervus nippon) in Japan
- japan_ferns_comm_full.csv: Community matrix (species x sites matrix) of native, non-hybrid ferns in Japan, full (unfiltered) dataset
- japan_ferns_inext_results.csv: Results of collection curve analysis on fern specimens in Japan
- japan_ferns_occ_summary.csv: Summary of specimen data of native, non-hybrid ferns in Japan
- japan_ferns_redundancy_by_res: Taxonomic richness, abundance, and redunancy at different spatial resolutions for native, non-hybrid ferns of Japan
- japan_ferns_shape_full.gpkg: Location of grid-cells (sites) for native, non-hybrid ferns in Japan, full (unfiltered) dataset
- japan_ferns_traits_lucid.csv: Morphological trait data of native, non-hybrid ferns in Japan
- japan_map_points.csv: Latitude and longitude of points used to produce a labeled map of Japan
- japan_map.gpkg: Map of Japan
- japan_protected_areas.gpkg: Protected areas in Japan
- testo_sundue_2016_calibrations.csv: Fossil calibration points used for phylogenetic dating analysis
- results.zip: Selected analysis results. Files include:
  - japan_ferns_biodiv.csv: Biodiversity statistics of native, non-hybrid ferns in Japan
  - japan_ferns_comm.csv: Community matrix (species x sites matrix) of native, non-hybrid ferns in Japan used for biodiversity analysis
  - japan_ferns_shape.gpkg: Location of grid-cells (sites) for native, non-hybrid ferns in Japan used for biodiversity analysis
  - japan_ferns_traits.csv: Trait matrix of native, non-hybrid ferns in Japan used for functional biodiversity analysis
  - japan_ferns_tree_dated.tre: Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in Japan with clades consisting of identical OTUs collapsed to polytomies
  - japan_ferns_tree_uncollapsed.tre: Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in Japan with clades consisting of identical OTUs not collapsed
  - japan_ferns_tree.tre: Maximim-likelihood phylogeny of native, non-hybrid ferns in Japan

@@

`r fs::path_file(deer_range_file)`: Distribution maps of Japanese deer (Cervus
nippon) in Japan. Raw data were downloaded from the Japan Ministry of the
Environment
(https://www.biodic.go.jp/biodiversity/activity/policy/map/map14/index.html).
The data include range maps based on three types of data: range of Japanese deer
surveyed in 1978, range surveyed in 2003, and estimated range inferred from a
model including snow cover and forest type based on the 2003 survey data. Only
estimated range with < 0.10 movement cost was included.

@@

MD5 checksum: `r tools::md5sum(here::here(inext_res_file))`

@@

```{r japan-ferns-biodiv-dryad}
japan_ferns_biodiv_dryad <- read_csv(here::here(japan_ferns_biodiv_dryad_file))
```

`r fs::path_file(japan_ferns_biodiv_dryad_file)` (contained in "results.zip"):
Biodiversity statistics of native, non-hybrid ferns and environmental variables
in Japan. Biodiversity metrics calculated as described in Nitta et al. 2021. 
Climatic (temperature and preciptation) variables
calculated as described for japan_climate.gpkg. Includes one row with missing
environmental data and one outlier for % apomixis that were removed prior to
spatial modeling analysis in Nitta et al. 2021.

@@

```{r japan-fern-dryad-uncollapsed-file}
japan_fern_tree_uncollapsed_dryad <- ape::read.tree(here::here(japan_fern_tree_uncollapsed_dryad_file))
```

`r fs::path_file(japan_fern_tree_uncollapsed_dryad_file)` (contained in "results.zip"):
Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in
Japan in newick format. Tree dated using treePL v1.0 [@Smith2012] with `r n_fossil_points` 
fossil calibration points after @Testo2016a. Values at nodes
are bootstrap support values calculated using 1000 replicates of ultrafast
bootstrap [@Nguyen2015]. Clades consisting of identical OTUs not collapsed.

Number of tips: `r ape::Ntip(japan_fern_tree_uncollapsed_dryad)`

MD5 checksum: `r tools::md5sum(here::here(japan_fern_tree_uncollapsed_dryad_file))`

@@

`r fs::path_file(japan_fern_tree_dryad_file)` (contained in "results.zip"):
Maximim-likelihood, ultrametric (dated) phylogeny of native, non-hybrid ferns in
Japan in newick format. Tree dated using treePL v1.0 [@Smith2012] with `r n_fossil_points` 
fossil calibration points after @Testo2016a. Values at nodes
are bootstrap support values calculated using 1000 replicates of ultrafast
bootstrap [@Nguyen2015]. Clades consisting of identical sequences have been
collapsed to polytomies with zero branch length between OTUs.

@@

2021-09-06
